\documentclass[../notes.tex]{subfiles}

\begin{document}

\subsection{Fields}
A field is said to exist when one object can exert a force on another object at a distance.

A gravitational field is associate with a mass.
Every mass has its own gravitational field.
A gravitational field only attracts, there is no repulsive gravitational field.

The gravitational field strength of an mass is $g$, or acceleration due to gravity.

\subsection{Field Lines}
Field lines help visualize and fields.
Field lines show the direction of the force due to a field.

Equipotential lines are drawn to represent where the potential of a field is always the same.
In a gravitational field line diagram, equipotential lines are farther apart from each other the farther the lines are from the center.

\section{Fields at Work}

\subsection{Gravitational Potential}
Gravitational potential difference is the work done in moving a unit mass between two points.
Gravitational potential is the work done in moving a unit mass infinitely far away to a certain point.
The equation for gravitational potential is:
\begin{equation}
	V_g = -\frac{GM}{r}
\end{equation}

Gravitational potential is always negative because the closer an object is to another object, the more work has been done to move the object from infinite distance, and thus the more work we will have to do to put it back.

\subsection{Gravitational Potential Energy}
Similar to gravitational potential, gravitational potential energy is the energy required to move a mass from infinite far away to a certain point.
\begin{equation}
	GPE = \frac{GMm}{r}
\end{equation}

\subsection{Escaping the Earth}
The total energy of a satellite is made up of its kinetic and gravitational potential energy.
In order for a satellite to escape a planet's gravitational field, the object will have to convert all of its potential energy into kinetic energy.
\begin{align}
	\Delta-\frac{GMm}{r} &= \Delta\frac{1}{2}mv^2
\end{align} 
Simplify this equation to find the escape velocity, which is $\sqrt{2gr}$.


\end{document}
